\newpage
\section*{ABSTRACT}
\addcontentsline{toc}{section}{ABSTRACT}
Contemporary conversational AI technologies are almost entirely based on the combination of Automatic Speech Recognition (ASR) and Large Language Models (LLM). Nevertheless, the majority of the current architectures embrace closely coupled ASR-LLM pipelines that constrain their flexibility, scalability and modularity. This project highlights a fully modular, bilingual speech-to-dialogue pipeline that decouples ASR, language modelling, retrieval, translation, and speech synthesis into separately optimizable parts which can be fine tuned to optimise each part independently without having to compromise the performance of others. The system leverages dedicated ASR models for English and Nepali speech input to deliver accurate bilingual transcription, translating Nepali output to English prior to language model inference. A lightweight Small Language Model (SLM) integrated with Retrieval-Augmented Generation (RAG) is then used to generate coherent, contextually relevant, and factually grounded responses. The system also supports audio output through bilingual English and Nepali Text-to-Speech modules, completing the full spoken dialogue pipeline. The English ASR model achieves a WER of 6.14\%, while the fine-tuned IndicConformer achieves a WER of 31.40\% on the clean Fleurs test set, outperforming all Whisper variants despite having fewer parameters, and 33.75\% under noisy conditions. The SLM achieves a perplexity of 1.96, a BERTScore-F1 of 0.9303, and a ROUGE-L of 0.5862, with RAG integration yielding significantly better factual grounding and relevance compared to the standalone model. This architecture is highly adaptable and well-performing in all its elements despite some of the trade-offs in terms of longer end-to-end latency, which validates the potential of a thoughtfully decoupled design in the evolution of scalable, maintainable, and multilingual conversational artificial intelligence systems.
\end{abstract}
\textit{Keywords: Automatic Speech Recognition (ASR), Parameter-Efficient Fine-Tuning (PEFT), Small Language Model (SLM), Retrieval-Augmented Generation (RAG), Text-to-Speech (TTS)} 
\pagebreak
